Molecular dynamics simulation of synchronization in driven particles

### 2020-11-06 manuscript/manuscript_ajp_2020.tex
updated text and section references

@@

%paper outline
In the following paper we describe
our molecular dynamics model in Section~\ref{sec:MD},
including various particle interactions and confining environments.
In Section~\ref{sec:sync} we
demonstrate how uniform environments and applied forces
and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams of particle velocity vs. position.
We modify the environment and applied forces
and show aperiodic, or nearly periodic flow behaviors
in Section~\ref{sec:quasiperiod}.
Finally we explore the transition to chaos in 
in Section~\ref{sec:chaos}.

@@

 We create several model light fields,
 creating a landscape of potential minima and maxima
 that modify the local force on a particle as a function of position
 $\vec{F}_{l,i} = \vec{F}_{l}(\vec{r}_i).$
 The landscape potential $V(\vec{r}) $ are static,
 generating a conservative force $\vec{F}_l= -\nabla V_l(\vec{r})$
 with fixed minima and maxima
 that are periodic or quasi-periodic,
 as described in Sec.~\ref{sec:results}.
 In every simulation, we confine
 the particles to a quasi-one-dimensional geometry
 using 
 \begin{equation}
   V_{q1D} = V_0 \cos{(\pi x / L)}
 \end{equation}.

 One example of a periodic light field is
 %that used in Sec.~\ref{sec:one}
 where 
 \begin{equation}
   V_l = V_0 \cos{(N_p \pi y / L)}
 \end{equation}
 where $N_p$ are the number of troughs in the landscape,
 and $V_0$ is an adjustable parameter
 to set the depth of the troughs.

@@

We place a single particle in a periodic light field is
\begin{equation}
  V_l = V_{0x} \cos{(\pi x/L)} + V_{0y} \cos{(N_p \pi y / L)} 
\end{equation}
where $N_p=20$ are the number of troughs in the landscape,
and $V_{0x}=2$ and $V_{0x}=2$.
This is illustrated in Fig.~\ref{fig:1}(a) where
the red regions show local maxima and the blue local minima.
We apply an external applied force, 
with 
a constant $F_{AC}$ with frequency $\omega$ then slowly increase $F_{DC}$.
The particle hops between troughs in the energy landscape
when the superposition of $F_{AC}$ and $F_{DC}$ is large enough
to overcome the barrier height.
By modifying $F_{DC}$ we 
achieve a variety of oscillation modes.
A mode is a periodic pattern of hops
with a constant average particle velocity, $\bar{v}_{y}$
over a range of driving forces $F_{DC}$.
We illustrate mode-locking in 
the velocity-force plot in Fig.~\ref{fig:1}(b).
Here $\bar{v}_{y}$ is increasing in non-uniform steps.
Our simulations reproduce results presented in 
Juniper {\it et al.} \cite{juniper2015}
which demonstrated
mode locking in
experiments of 
driven colloids on a
optical periodic landscape.

@@

\section{Associated problems}

\section{Conclusion}

@@

We acknowledge Harvey Gould and Jan Tobochnik, who supported our development of this article.  We acknowledge funding from the M.J. Murdock Charitable Trust and the Pacific PRISM program.

>>>>

%another confinement technique:
Acoustic tweezers and surface acoustic wave (SAW) \cite{}
%confines single cells and particles with lower energy than light / lasers
%%X. Ding, J. Shi, S.-C. S. Lin, S. Yazdi, B. Kiraly and T. J. Huang, “Tunable patterning of microparticles and cells using standing surface acoustic waves” Lab Chip, 2012, 12, 2491– 2497.
%%%Acousto-optically generated potential energy landscapes: Potential mapping using colloids under flow Michael P. N. Juniper,1,∗ Rut Besseling,2 Dirk G. A. L. Aarts,1 and Roel P. A. Dullens1

@@

%----------------------------------------------------------------------------
%paper outline
In the following paper we describe
our molecular dynamics model in Section~\ref{sec:MD}.
%including various  and confining environments.
We include code to simulate
and visualize the dynamics in this section
and supplementary material.
In Section~\ref{sec:results} we summarize
our results,
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one} and
multiple interacting particles
in Section~\ref{sec:sync}.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams of particle velocity versus position.
In Section~\ref{sec:quasiperiod},
we show how an aperiodic landscape modifies the particle dynamics.
Finally we explore the transition to chaotic dynamics in 
in Section~\ref{sec:chaos}.
Our conclusions
are in Section~\ref{sec:conclusion}. 
In each section
we suggest exercises for interested students,
and summarize our suggestions in Section~\ref{sec:problems}.

@@

 We create several model light fields,
 creating a landscape of potential minima and maxima
 that modify the local force on a particle as a function of position
 $\vec{F}_{l,i} = \vec{F}_{l}(\vec{r}_i).$
 The landscape potential $V(\vec{r}) $ are static,
 generating a conservative force $\vec{F}_l= -\nabla V_l(\vec{r})$
 with fixed minima and maxima
 that are periodic or quasi-periodic,
 as described in Sec.~\ref{sec:results}. 
 One example of a periodic light field is
 %that used in Sec.~\ref{sec:one}
 where 
 \begin{equation}
   V_l = V_0 \cos{(N_p \pi y / L)}
 \end{equation}
 where $N_p$ are the number of troughs in the landscape,
 and $V_0$ is an adjustable parameter
 to set the depth of the troughs.

 In multi-particle simulations, we confine
 the particles along the x-direction with a
 quasi-one-dimensional geometry
 using a periodic function.
 \begin{equation}
   V_{q1D} = V_0 \cos{(\pi x / L)}
 \end{equation}.
 %%We could also use a Gaussian function.

@@

We place a single particle in a periodic landscape 
\begin{equation}
  \label{eq:landscape}
  V_l = V_{0x} \cos{(\pi x/L)} + V_{0y} \cos{(N_p \pi y / L)} 
\end{equation}
where $N_p=20$ are the number of troughs in the landscape,
and $V_{0x}=2$ and $V_{0x}=2$.
This is illustrated in Fig.~\ref{fig:1}(a) where
the red regions show local maxima and the blue local minima.
The code for generating
a two dimensional colored plot
of the potential landscape
is calculated by evaluating
the analytic function in Eq.~\ref{eq:landscape}
for a grid of values $(x_n,y_n)$.
%is included in the supplementary material,

We apply an external applied force, 
with 
a constant $F_{AC}$ with frequency $\omega$ then slowly increase $F_{DC}$.
The particle hops between troughs in the energy landscape
when the superposition of $F_{AC}$ and $F_{DC}$ is large enough
to overcome the barrier height.
By modifying $F_{DC}$ we 
achieve a variety of oscillation modes.
A mode is a periodic pattern of hops
with a constant average particle velocity, $\bar{v}_{y}$
over a range of driving forces $F_{DC}$.
We illustrate mode-locking in 
the velocity-force plot in Fig.~\ref{fig:1}(b).
Here $\bar{v}_{y}$ is increasing in non-uniform steps.
Our simulations reproduce results presented in 
Juniper {\it et al.} \cite{juniper2015}
which demonstrated
mode locking in
experiments of 
driven colloids on a
optical periodic landscape.

@@

\section{Associated problems}
\label{sec:problems}	% You can label sections for reference

\section{Conclusion}
\label{sec:conclusion}	% You can label sections for reference

@@

  We acknowledge Harvey Gould and Jan Tobochnik, who supported our development of this article.
  Charles and Cynthia Reichhardt advised 
  the project and provided the molecular dynamics code.
  We acknowledge funding from the M.J. Murdock Charitable Trust
  and the Pacific Research Institute for Science and Mathematics (PRISM).
